\bfseries A simple, anisotropic ram-pressure model for sailing

### 2021-06-21 paper/main.tex
lots of figures, figure captions, and text

% style notes:
% ------------
% - Always air before water, always sail before keel before hull.

% to-do items:
% ------------
% - FIX THE FORCE EXPRESSION and discuss it in terms of inner products, tensors, and Einstein summation notation.
% - Put in running head.
% - Overleaf seems to have wrong papersize?
% — Put in the dumbell figure
% - Reconsider the post-equation commas? (HWR SAY).
% — Find references for, and destroy, Bernoulli.
% — Make some comments on the situations in which the ram-pressure force approximation is more appropriate---gale force or lull?

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{{./}{../ipynb/}}

@@

% text macros
\usepackage{biblatex}
\addbibresource{sailing.bib}
\newcommand{\documentname}{\textsl{Article}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}

@@

HOGG: We make use of good, coordinate-free expressions.
We are Galilean-relativistic.

@@

Finally, although upwind travel is slow going, with lots of tacking (HOGG DO WE NEED TO DEFINE THIS?) required, if you are situated \emph{on the boat}, upwind travel \emph{feels} very fast.
Why?
Because your experience of speed is, at least in part, the feeling of the wind in your face.
If you think about \figref{fig:hourglass} in terms of the boat-air velocity instead of the boat-water velocity, the upwind travel has much larger boat-air velocities than the downwind travel.
So even though you can sail fast downwind, it sometimes \emph{feels} faster upwind.

@@

HOGG: FIGURE showing good sailing for a set of boat orientations.
\begin{figure}[t!]
  \includegraphics[width=\textwidth]{good.pdf}
  \caption{Visualizations of GoodSailing(tm) for different settings of the orientation of the keel.
  Conventions are the same as in \figref{fig:steady}.
  .\label{fig:good}}
  \figurerule
\end{figure}

HOGG: Announce and discuss \figref{fig:hourgood}.
\begin{figure}[t!]
  ~\hfill\includegraphics[width=3in]{hourglass-good.pdf}\hfill~
  \caption{The $x$ and $y$ coordinates of the relative boat--water velocity $\vboat-\vwater$ for GoodSailing(tm) for a set of randomly oriented keels.
  For all of these trials, the relative wind--water velocity is $\vair-\vwater=(8\,\mps)\,\uvec_x$ (that is, $8\,\mps$ in the $x$ direction), as it is in every panel of \figref{fig:good}.
  Compare this to \figref{fig:hourglass}.\label{fig:hourgood}}
  \figurerule
\end{figure}

@@

A type-A sailor---or maybe a competetive sailor---is trying to get from a current position to a destination as quickly as possible\footnote{%
We don't particularly endorse this attitude towards sailing. The journey \emph{is} the destination.}.
Our view is that this sailor should start by specifying the unit vector $\uvec_\destination$ that points from the current position of the boat towards the destination of the boat.
The sailor should then maximize the scalar product $\uvec_\destination^\top\,\vboat$
\begin{align}\label{eq:better}
    \uvec_{\perp\sail}^\better,\uvec_{\perp\keel}^\better &\leftarrow \argmax_{\uvec_{\perp\sail},\uvec_{\perp\keel}} \left[\uvec_\destination^\top\,(\vboat-\vdest)\right] ~,
\end{align}
where $\uvec_{\perp\sail}^\better,\uvec_{\perp\keel}^\better$ are the best settings of the orientations of the keel and sail, given the direction $\uvec_\destination$ to the destination, and (because we are Galilean-relativistic) this velocity is relative to the velocity $\vdest$ of the destination (which might not be stationary with respect to the water; imagine that we are sailing on a fast-flowing river).
In this expression \eqref{eq:better}, implicitly $\vboat$ is being thought of as a function of the orientations (and more).
Note that the sailor \emph{does not necessarily want} to travel directly towards the destination:
The sailor wants to make as much progress as possible in that direction, but is willing to tack back-and-forth to get there, if it gets the boat there faster.
So the metric really is the projection of the boat velocity onto the displacement vector pointing from the current position of the boat to the destination.

While this is simple to state, the numerical optimization hurts...HOGG

HOGG PUT FIGURES HERE SHOWING THE OPTIMAL SETTINGS OF SAIL AND KEEL ORIENTATIONS FOR DIFFERENT DESTINATION DIRECTIONS.

@@

HOGG: These properties might seem unreasonable! However, remember that our boat is just a brutal approximation.
A true boat has sails that are globally curved\footnote{%
In the local-curvature or Riemann-curvature senses, sails are not (very) curved; after all, they are made from flat pieces of fabric.
However, they are curved in the global or colloquial sense: Different patches of the sail have differently oriented normal vectors.},
and forces that are different from these pure ram-pressure forces.
The curvature of the sails will in general rotate the air force vectors towards the direction in which the boat is pointed.

@@

HOGG: How we implement GoodSailing(tm) and BetterSailing(tm). It's a mess!

>>>>

% style notes:
% ------------
% - Always air before water, always sail before keel before hull.
% - Need to choose typesetting for GoodSailing and BetterSailing and then be consistent.

% to-do items:
% ------------
% - FIND THE PRIOR LITERATURE!!
% - Overleaf seems to have wrong papersize?
% - Reconsider the post-equation commas? (HWR SAY).
% — Find references for, and destroy, Bernoulli.
% — Make some comments on the situations in which the ram-pressure force approximation is more appropriate---gale force or lull?
% - Is the latex footnote-footnote separation fucked?

\documentclass[letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{{./}{../ipynb/}}
\usepackage{biblatex}
\addbibresource{sailing.bib}

@@

% text macros
\newcommand{\documentname}{\textsl{Article}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}

@@

HOGG: We make use of good, coordinate-free expressions, a la \cite{kusse}.
We are Galilean-relativistic.

@@

Finally, although upwind travel is slow going, with lots of tacking (HOGG DO WE NEED TO DEFINE THIS?) required, if you are situated \emph{on the boat}, upwind travel \emph{feels very fast}.
Why?
Because your experience of speed is, at least in part, the feeling of the wind in your face.
If you think about \figref{fig:hourglass} in terms of the boat-air velocity instead of the boat-water velocity---that is, if you Galilean-boost to the rest frame of the air---the upwind travel has much larger boat-air velocities than the downwind travel.
So even though you can sail fast downwind, it sometimes \emph{feels} faster upwind.

@@

HOGG: FIGURE showing good sailing for a set of boat orientations.
\begin{figure}[t!]
  \includegraphics[width=\textwidth]{good.pdf}
  \caption{Visualizations of GoodSailing(tm) for different settings of the orientation of the keel.
  Conventions are the same as in \figref{fig:steady}.
  GoodSailing(tm) is unsuccessful in the last two cases, where the pointing is too close to directly upwind.
  .\label{fig:good}}
  \figurerule
\end{figure}

HOGG: Announce and discuss \figref{fig:hourgood}.
\begin{figure}[t!]
  ~\hfill\includegraphics[width=3in]{hourglass-good.pdf}\hfill~
  \caption{The $x$ and $y$ coordinates of the relative boat--water velocity $\vboat-\vwater$ for GoodSailing(tm) for a set of randomly oriented keels.
  For all of these trials, the relative wind--water velocity is $\vair-\vwater=(8\,\mps)\,\uvec_x$ (that is, $8\,\mps$ in the $x$ direction), as it is in every panel of \figref{fig:good}.
  Compare this to \figref{fig:hourglass}.
  As in \figref{fig:good}, when the keel orientation is too close to directly upwind, the boat goes downwind instead.\label{fig:hourgood}}
  \figurerule
\end{figure}

@@

A type-A sailor---or maybe a competetive sailor---is trying to get from a current position to a destination as quickly as possible\footnote{%
We don't particularly endorse this attitude towards sailing. The journey \emph{is} the destination.}.
Our view is that this sailor should start by specifying the unit vector $\uvec_\destination$ that points from the current position of the boat towards the destination of the boat.
The sailor should then maximize the scalar product $\uvec_\destination^\top\,\vboat$
\begin{align}\label{eq:better}
    \uvec_{\perp\sail}^\better,\uvec_{\perp\keel}^\better &\leftarrow \argmax_{\uvec_{\perp\sail},\uvec_{\perp\keel}} \left[\uvec_\destination^\top\,(\vboat-\vdest)\right] ~,
\end{align}
where $\uvec_{\perp\sail}^\better,\uvec_{\perp\keel}^\better$ are the best settings of the orientations of the keel and sail, given the direction $\uvec_\destination$ to the destination, and (because we are Galilean-relativistic) this velocity is relative to the velocity $\vdest$ of the destination (which might not be stationary with respect to the water; imagine that we are sailing on a fast-flowing river).
In this expression \eqref{eq:better}, implicitly $\vboat$ is being thought of as a function of the orientations (and more).
Note that the sailor \emph{does not necessarily want} to travel directly towards the destination:
The sailor wants to make as much progress as possible in that direction, but is willing to tack back-and-forth to get there, if it gets the boat there faster.
So the optimization objective really is the projection of the boat velocity onto the displacement vector pointing from the current position of the boat to the destination.
While this is all simple to state, the global optimization is a bit painful to implement in code.
We give details in \secref{sec:implementation}.

HOGG PUT FIGURES HERE SHOWING THE OPTIMAL SETTINGS OF SAIL AND KEEL ORIENTATIONS FOR DIFFERENT DESTINATION DIRECTIONS.
\begin{figure}[t!]
  \includegraphics[width=\textwidth]{better.pdf}
  \caption{Visualizations of BetterSailing(tm) for different directions to a distant destination.
  Conventions are the same as in \figref{fig:steady} and \figref{fig:good}, except that now the destination direction is indicated with a dashed path line.
  BetterSailing(tm) often involves sailing at a substantial angle to the direct line towards the destination.\label{fig:better}}
  \figurerule
\end{figure}

@@

HOGG: These properties might seem unreasonable! However, remember that our boat is just a brutal approximation.
A true boat has sails that are globally curved\footnote{%
In the local-curvature or Riemann-curvature senses, sails are not (very) curved; after all, they are made from flat pieces of fabric.
However, they are curved in the global or colloquial sense: Different patches of the sail have differently oriented normal vectors.
It is in this colloquial sense that we are using the words ``curved'' and ``flat''.},
and forces that are different from these pure ram-pressure forces.
The curvature of the sails will in general rotate the air force vectors towards the direction in which the boat is pointed.

@@

HOGG: How we implement GoodSailing(tm) and BetterSailing(tm) optimizations. It's a mess!
